A Formalization of Complete Discrete Valuation Rings and Local Fields

### 2023-12-11 HAL/CPP_paper_v2.tex
update intro

From a geometric perspective, one can regard the previous as a kind of ``local analysis'': interpreting the primes as the points of a geometric object attached to $\ZZ$, if a ``global'' solution $(a_1,a_2,a_3)\in\ZZ^3$ exists, then for each prime $p$, we obtain a solution $(\overline{a_1},\overline{a_2},\overline{a_3})\in\FF^3$ by reducing the coordinates modulo $p$ --- we think of this as a local solution ``around the point $p$'' --- as well as a solution  $((a_1 : \RR),(a_2 : \RR),(a_3 : \RR))\in\RR^3$ over the reals. \mi{Conversely}, if one of these local solutions fails to exists, this means that the equation cannot have a global solution. Yet, in this analogy, the fields $\FF$ and $\RR$ are very different: the former is finite, with trivial discrete geometry and of positive characteristic, while the latter has a rich metric structure and characteristic $0$. Now, for every $p$, one can rather consider the field of $p$-adic numbers $\QQ_p$ which is, in this perspective, a better analogue for $\RR$ than the several $\FF$: it has a metric with respect to which it is complete (Cauchy sequences converge), and it has characteristic $0$ (so, in particular, it contains $\QQ$). To define it, one observes that for every prime $p$ there exists an absolute value $\lvert\cdot\rvert_p$ on $\QQ$ for which numbers that are highly divisible by $p$ are close to $0$. Having an absolute value, it is possible to speak about convergence and about Cauchy sequences: exactly as for the euclidean absolute value, one can find Cauchy sequences that do not converge to any rational, and $\QQ_p$ is defined as the \emph{completion} of $\QQ$ with respect to~$\lvert\cdot\rvert_p$. It is the ``smallest'' field where all $\lvert\cdot\rvert_p$-Cauchy sequences converge. Crucially, it still bears strong connections with the prime $p$: for instance, if a monic polynomial in $\ZZ[X]$ has a simple root in $\FF$, then it also has one in $\QQ_p$.

The local analysis described before is an example of the ``local-to-global'' principle: to solve a problem in $\ZZ$, or $\QQ$, one can try to solve it in $\QQ_p$, for all $p$, and in $\RR$. A famous application of the local-to-global principle is \mi{found in} the Hasse–Minkowski Theorem, which states that if $Q(x)$ is a quadratic form with rational coefficients and $c \in \unit{\QQ}$, then the equation $Q(x) =c$ is solvable over $\QQ$ if and only if it is solvable over $\RR$ and over every $\QQ_p$.

@@

For an in-depth discussion of bundled versus unbundled representations \fil{in Lean and in Coq}, we refer the reader to \cite{Baa22} or \cite{typeclassesCoq}.

@@

\begin{acks}
We would like to thank Riccardo Brasca, Mario Carneiro, Antoine Chambert-Loir and Rob Lewis for interesting discussions about possible formalizations of the $p$-adic numbers, Heather Macbeth for helpful conversations about the formalization of Laurent series, Yaël Dillies for providing access to some results about well-founded relations\ \fil{and Cyril Cohen for answering some questions about $p$-adic numbers in Coq}. We also thank the \mathlib community as a whole for their support during the completion of this project.

>>>>

From a geometric perspective, one can regard the previous as a kind of ``local analysis'': interpreting the primes as the points of a geometric object attached to $\ZZ$, if a ``global'' solution $(a_1,a_2,a_3)\in\ZZ^3$ exists, then for each prime $p$, we obtain a solution $(\overline{a_1},\overline{a_2},\overline{a_3})\in\FF^3$ by reducing the coordinates modulo $p$ --- we think of this as a local solution ``around the point $p$'' --- as well as a solution  $((a_1 : \RR),(a_2 : \RR),(a_3 : \RR))\in\RR^3$ over the reals. In particular, if one of these local solutions fails to exists, this means that the equation cannot have a global solution. Yet, in this analogy, the fields $\FF$ and $\RR$ are very different: the former is finite, with trivial discrete geometry and of positive characteristic, while the latter has a rich metric structure and characteristic $0$. Now, for every $p$, one can rather consider the field of $p$-adic numbers $\QQ_p$ which is, in this perspective, a better analogue for $\RR$ than the several $\FF$: it has a metric with respect to which it is complete (Cauchy sequences converge), and it has characteristic $0$ (so, in particular, it contains $\QQ$). To define it, one observes that for every prime $p$ there exists an absolute value $\lvert\cdot\rvert_p$ on $\QQ$ for which numbers that are highly divisible by $p$ are close to $0$. Having an absolute value, it is possible to speak about convergence and about Cauchy sequences: exactly as for the euclidean absolute value, one can find Cauchy sequences that do not converge to any rational, and $\QQ_p$ is defined as the \emph{completion} of $\QQ$ with respect to~$\lvert\cdot\rvert_p$. It is the ``smallest'' field where all $\lvert\cdot\rvert_p$-Cauchy sequences converge. Crucially, it still bears strong connections with the prime $p$: for instance, if a monic polynomial in $\ZZ[X]$ has a simple root in $\FF$, then it also has one in $\QQ_p$.

The local analysis described before is an example of the ``local-to-global'' principle: to solve a problem in $\ZZ$, or $\QQ$, one can try to solve it in $\QQ_p$, for all $p$, and in $\RR$. A famous application of the local-to-global principle is the Hasse–Minkowski Theorem, which states that if $Q(x)$ is a quadratic form with rational coefficients and $c \in \unit{\QQ}$, then the equation $Q(x) =c$ is solvable over $\QQ$ if and only if it is solvable over $\RR$ and over every $\QQ_p$.

@@

For an in-depth discussion of bundled versus unbundled representations in Lean and in Coq, we refer the reader to \cite{Baa22} or \cite{typeclassesCoq}.

@@

\begin{acks}
We would like to thank Riccardo Brasca, Mario Carneiro, Antoine Chambert-Loir and Rob Lewis for interesting discussions about possible formalizations of the $p$-adic numbers, Heather Macbeth for helpful conversations about the formalization of Laurent series, Yaël Dillies for providing access to some results about well-founded relations and Cyril Cohen for answering some questions about $p$-adic numbers in Coq. We also thank the \mathlib community as a whole for their support during the completion of this project.
